Large-scale harmonic patterns clustering

### 2010-03-20 Paper/Paper.tex
improved

\section{Algorithm}
Wth the scaling factor as our main concern, we present an online
vector quantization algorithm to cluster the harmonic patterns.

@@

>>>>

\section{Algorithm}\label{sec:algo}
In this section we present the creation of the harmonic patterns from
the Echo Nest data. Then, we explain the online vector quantization algorithm
used to learn a codebook of those patterns.

\subsection{Pattern Creation}
Here we should explain key invariance, resizing, etc, and not in the
experiment section.

We try to normalize the patterns in respect to the key. We roll the matrix
so the row with the more energy is always the lowest one. This can easily be
seen in Figure \ref{fig:code}. We considered the same technique for rolling
along the other axis and be invariant regarding the downbeat. However, it
seems less robust to noise. Note that we roll each pattern independentely.
Also, this implies that the pitch envelope is important rather than the actual 
pitches.

\subsection{Codebook Learning}
We use an online version of the vector quantization algorithm 
\cite{Gersho1991}. This can also be seen as an online K-means.
The idea is as follow: take a sample from your data, find the closest
code in the codebook, bring that code closer to the sample by some amount.
The details are explained in Algorithm \ref{algo:vq}. Note that this
algorithm, though not optimal, is $O(n)$ with $n$ number of songs seen.
Thus it scales linearly to any size of data set. The setting of the learning
rate might become trickier, but this is not fundamentally a problem.

We initialize the codebook by choosing $K$ random points from our dataset.

@@

\subsection{Setting}\label{ssec:setting}
We take one or two bars, normalize the patches to size 4, 8, or 16.
We roll the patch to be invariant to the key (on the patch level, not on
the song level). We learn a codebook of size $K$ over the cowbell dataset 
using the VQ algorithm (Algorithm \ref{algo:vq}). A typical learning rate 
is $0.01$ for $200$ iterations over the whole dataset.

We use the codebook to encode our testing set (see Section \ref{sec:data}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes, i.e. the size of the cluster around a code.

As a distance measure (or distortion measure) between patterns, we use
the average squared euclidean distance. In details, if a pattern $p1$
is composed of elements $p1_{i,j}$, and similarly for a pattern $p2$,
the distance between $p1$ and $p2$ is:
\begin{eqnarray}
  dist(p1,p2) = \sum_{i,j} (p1_{i,j} - p2_{i,j})^2 / size(p1)  \label{eq:dist}
\end{eqnarray}
We assume $p1$ and $p2$ have the same size.
